---
title: "Predicting Playoffs from Four Factors using Logistic Regression"
author: "Matthew Wankiewicz"
date: "24/10/2020"
abstract: ""
output: html_document
---

```{r}
library(tidyverse)
library(InformationValue)
library(rvest)
```
# Introduction

# Data

```{r}
nba_data <- read_csv("nba_data.csv")
nba_data <- nba_data %>% 
  mutate(team_name = paste(TEAM, year, sep = " "))
```

```{r}
nba_data %>% 
  ggplot(aes(W, efg, color = playoffs)) +
  geom_point()
```

```{r}
nba_data %>% 
  ggplot(aes(W, fta, color = playoffs)) +
  geom_point()
```

```{r}
nba_data %>% 
  ggplot(aes(W, tov, color = playoffs)) +
  geom_point()
```

```{r}
nba_data %>% 
  ggplot(aes(W, oreb, color = playoffs)) +
  geom_point()
```

```{r}
nba_data %>% 
  ggplot(aes(W, efg_o, color = playoffs)) +
  geom_point()
```

```{r}
nba_data %>% 
  ggplot(aes(W, fta_o, color = playoffs)) +
  geom_point()
```

```{r}
input_ones <- nba_data[which(nba_data$playoffs == 1), ]  # all 1's
input_zeros <- nba_data[which(nba_data$playoffs == 0), ]  # all 0's
set.seed(21)  # for repeatability of samples
input_ones_training_rows <- sample(1:nrow(input_ones), 0.7*nrow(input_ones))  # 1's for training
input_zeros_training_rows <- sample(1:nrow(input_zeros), 0.7*nrow(input_ones))  # 0's for training. Pick as many 0's as 1's
training_ones <- input_ones[input_ones_training_rows, ]  
training_zeros <- input_zeros[input_zeros_training_rows, ]
trainingData <- rbind(training_ones, training_zeros)  # row bind the 1's and 0's 

# Create Test Data
test_ones <- input_ones[-input_ones_training_rows, ]
test_zeros <- input_zeros[-input_zeros_training_rows, ]
testData <- rbind(test_ones, test_zeros) 
```


```{r}
logitMod <- glm(playoffs ~ efg + efg_o + oreb + oreb_o +
                  fta + fta_o + tov + tov_o, 
                data=trainingData, family=binomial(link="logit"))
summary(logitMod)
predicted <- predict(logitMod, testData, type="response")
```

```{r}
nba_data$estimate <- predict.glm(logitMod,
                                 newdata = nba_data,
                                 type = "response")

knitr::kable(nba_data, digits = 2) %>% kableExtra::kable_styling()
```



```{r}
optCutOff <- optimalCutoff(testData$playoffs, predicted)[1] 

plotROC(testData$playoffs, predicted)
sensitivity(testData$playoffs, predicted, threshold = optCutOff)
confusionMatrix(testData$playoffs, predicted, threshold = optCutOff)
```


```{r}
new_df %>% ggplot(aes(playoff_prob, playoffs)) + geom_point()
```

## References

- R Core Team (2019). R: A language and environment for statistical computing. R Foundation for
  Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
  
- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43),
  1686, https://doi.org/10.21105/joss.01686
  
- NBA (2020), Advanced Stats,
  https://stats.nba.com/teams/four-factors/?sort=W&dir=-1&Season=2010-11&SeasonType=Regular%20Season
  
- Selva Prabhakaran (2016). InformationValue: Performance Analysis and Companion Functions for
  Binary Classification Models. R package version 1.2.3.
  https://CRAN.R-project.org/package=InformationValue
